\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Madrid}
  \usecolortheme{rose}
  \usefonttheme{structurebold}
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{listings}
\usepackage{inconsolata}
\usepackage{qrcode}

\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}

\usepackage{upquote}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=fullflexible,
  frame=single,
  showstringspaces=false
}

\title[Mill as an Alternative Android Build Tool]{Until recently, Gradle was the only realistic option for Android builds.}
\author{Vasilis Nicolaou}
\institute{Vaslabs LTD}
\date{Android x GDG Manchester winter meetup @ 22 January 2026}

\begin{document}

% =========================
% Title & Agenda
% =========================

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

% =========================
% Intro / Context
% =========================

\section{Introduction}

\begin{frame}{Who am I?}
  \begin{itemize}
    \item JVM developer (Java, Scala) with experience in distributed systems.
    \item Contributor to Mill: Android, GraalVM, Spring Boot, Micronaut.
    \item I went through various build tools; Maven, Gradle, Sbt and now... Mill!
    \item Today: Software Engineering Consultant with own firm and a small but strong team of software engineers!
  \end{itemize}
\end{frame}

% =========================
% Live Demos First
% =========================

\section{Mill Live Demo}

\subsection{Architecture Samples (Hilt/KSP, Simple Compose)}

\begin{frame}{Architecture Samples (BuildConfig, Hilt/KSP, Simple Compose)}
  \vspace{0.5em}
  \textbf{Basic CLI usage:}
  \begin{lstlisting}[language=bash, basicstyle=\ttfamily\tiny]
./mill app.compile
./mill app.androidApk
./mill app.androidTest.androidTestApk
./mill app.createAndroidVirtualDevice
./mill app.startAndroidEmulator
./mill app.androidInstall
./mill app.androidRun --activity com.example.android.architecture.blueprints.todoapp.TodoActivity
./mill app.androidTest
./mill app.stopAndroidEmulator
\end{lstlisting}
\end{frame}

\subsection{Compose Samples}


\begin{frame}{JetNews}
  \vspace{0.5em}
  \textbf{Basic CLI usage:}
  \begin{lstlisting}[language=bash, basicstyle=\ttfamily\tiny]
./mill JetNews.app.androidApk
./mill JetNews.app.androidTest.androidTestApk
./mill app.startAndroidEmulator
./mill app.androidInstall
./mill app.androidTest
\end{lstlisting}
\end{frame}

\begin{frame}{Jetcaster}
  \vspace{0.5em}
  \textbf{Basic CLI usage:}
  \begin{lstlisting}[language=bash, basicstyle=\ttfamily\tiny]
./mill Jetcaster.wear.androidApk
./mill Jetcaster.wear.androidTest.androidTestApk
./mill Jetcaster.wear.startAndroidEmulator
./mill Jetcaster.wear.androidInstall
./mill Jetcaster.wear.androidTest
\end{lstlisting}
\end{frame}

% =========================
% Comparison & Wrap-up
% =========================

\section{Closing Remarks}

\subsection{Key Takeaways}

\begin{frame}{Key Takeaways}
  \begin{itemize}
    \item Mill treats builds as real code, not opaque configuration (but if you wish, you can also use the latter with YAML (new)).
    \item The task graph is explicit, inspectable, and composable.
    \item Mill can support modern JVM apps with AOT and native configurations.
    \item Android support demonstrates that Mill scales to highly complex build pipelines.
    \item You can take control of your build by easily understanding and extending it.
    \item Rich examples are documented for a variety of use cases and always validated and tested against each change in Mill.
  \end{itemize}
  \begin{block}{Summary}
    Mill is a Swiss knife. But there are efforts to have also the option for non programmatic usage if one wishes to do so!
  \end{block}
\end{frame}

\subsection{Q\&A}

\begin{frame}{Questions}
  \begin{itemize}
    \item Which version of Spring boot are you using?
    \item What is your library stack? (Mono/Flux, rxJava, Lombok, Immutables)
    \item Do you deploy native images? Are you using AOT?
    \item Any other testing libraries/pre-processors?
    \item What is the biggest pain on the build tool you are using?
    \item Plain Java? Kotlin? Mixed?
  \end{itemize}
\end{frame}

\begin{frame}{Thank you!}
  \centering
  \Large{Thank you!}\\[1em]
  \normalsize
  \begin{itemize}
      \item Documentation: \url{https://mill-build.org/}
      \item Github Repo: \url{https://github.com/com-lihaoyi/mill}
      \item Try Mill and join Mill's Discord Channel:
  \end{itemize}

  \quad
  \qrcode[height=5cm]{https://discord.gg/bt3DPbfUmV}

\end{frame}

\end{document}
